Bootstrapping Domain Specific Languages for Neurally-Guided Bayesian Program Learning

### 2018-02-07 paper.tex
shrink font size of code

\newcommand{\system}{\textsc{DreamCoder}~}
\newcommand{\lowerBound}{\mathscr{L}}
% Use the following line for the initial blind version submitted for review:
\usepackage[nohyperref]{icml2018}

@@

We apply \system to four domains:
 building Boolean circuits; symbolic regression; FlashFill-style~\cite{gulwani2011automating} string processing problems; and Lisp-style functions on lists.
 For each of these we deliberately provide an impoverished
 set of programming primitives,
 and show that our algorithm discovers
 its own domain-specific vocabulary for expressing solutions the domain (Table~\ref{initialExampleDSL}).
 \begin{table}
    \begin{tabular}{ll}
   \toprule
   Domain&Part of the learned DSL\\\midrule
   Circuits&\texttt{(nand (nand a a) (nand b b))}\\
   &\hspace{0.5cm} \emph{(an OR gate)}\\
   Regression& \texttt{(+ (* real x) real)} \\
      &\hspace{0.5cm} \emph{(a linear function of x)}\\
   Strings& \texttt{(join '' (split c s))}\\
   &\hspace{0.5cm} \emph{(delete occurrences of c in string s)}\\
   Lists& \texttt{(map (lambda (x) (+ x a)) l)}\\
   &\hspace{0.5cm} \emph{(add a to every element of list l)}
   \\\bottomrule
    \end{tabular}
    \caption{ Examples of structure found in DSLs  learned by our algorithm. \system builds a new DSL by discovering and reusing useful subroutines.}\label{initialExampleDSL}
 \end{table}

@@

To make this concrete we need to define what programs actually are and
what form $\probability[p |\mathcal{D},\theta]$ takes.
We represent programs as $\lambda$-calculus expressions.
$\lambda$-calculus is a formalism for expressing functional programs
that closely resembles the Lisp programming language.
$\lambda$-calculus includes variables, function application, and the ability to create new functions.
Throughout this paper we will write $\lambda$-calculus expressions in Lisp syntax.
Our programs are all strongly typed.
We use the Hindley-Milner polymorphic typing system~\cite{pierce} which is used in functional languages like OCaml.
Variables that range over types are always written using lowercase Greek letters
and we write $\alpha\to \beta$ to mean a function that takes as input type $\alpha$
and returns something of type $\beta$.
We use the notation $p:\tau$ to mean that the $\lambda$-calculus expression $p$
has the type $\tau$.
For example, to describe the type of the identity function
we would say \texttt{(lambda (x) x)}$:\alpha\to \alpha$.
We say a type $\alpha$ \emph{unifies} with $\tau$ if every expression
$p:\alpha$ also satisfies $p:\tau$. Furthermore, the act of \emph{unifying}
a type $\alpha$ with $\tau$ is to introduce constraints on the type
variables of $\alpha$ to ensure that $\alpha$ unifies with $\tau$.
See Supplement for more detail on program representation.
With this notation in hand we can now define DSLs:

@@

\begin{algorithm}[tb]
   \caption{Generative model over programs}
   \label{programGenerativeModel}
   \begin{algorithmic}
     \STATE \textbf{function} sampleProgramFromDSL$(\mathcal{D}, \theta, \tau)$:
  \STATE {\bfseries Input:} DSL $\mathcal{D}$, weight vector $\theta$, type $\tau$
  \STATE \textbf{Output:} a program whose type unifies with $\tau$
  \STATE \textbf{return} sample$(\mathcal{D}, \theta, \varnothing, \tau)$
\STATE
     \STATE \textbf{function} sample$(\mathcal{D}, \theta, \mathcal{E}, \tau)$:
  \STATE {\bfseries Input:} DSL $\mathcal{D}$, weight vector $\theta$, environment $\mathcal{E}$, type $\tau$
  \STATE \textbf{Output:} a program whose type unifies with $\tau$
  \IF{$\tau = \alpha\to\beta$}
  \STATE var $\gets$ an unused variable name
  \STATE body $\sim$ sample$(\mathcal{D},\theta,\{\text{var}:\alpha\}\cup\mathcal{E},\beta)$
   \STATE \textbf{return} \texttt{(lambda (}var\texttt{) }body\texttt{)}
   \ENDIF
   %   \ELSE
   \STATE $\text{primitives} \gets\{p | &p: \tau' \in \mathcal{D}\cup\mathcal{E}$
   \STATE \hspace{2.5cm}$\text{if }\tau\text{ can unify with yield}(\tau') \} $

   \STATE Draw $e\sim \text{primitives}$, w.p. $\propto\theta_e$ if $e\in \mathcal{D}$
   \STATE \hspace{3.1cm}w.p. $\propto\frac{\theta_{var}}{|\text{variables}|}$ if $e\in \mathcal{E}$
   \STATE Unify $\tau$ with yield$(\tau')$.
   \STATE $\left\{\alpha_k \right\}_{k\leq K}\gets\text{argTypes}(\tau')$
%   \STATE unify$(\tau,\beta)$
   \FOR{$k=1$ {\bfseries to} $K$}
 \STATE $a_k\sim\text{sample}(\mathcal{D},\theta,\mathcal{E},\alpha_k)$
 \ENDFOR
 \STATE \textbf{return} \texttt{(}$e\;a_1\; a_2\; \cdots\; a_K$\texttt{)}
 \STATE\textbf{where:}
 \STATE  yield$(\tau) = \begin{cases}
\text{yield}(\beta)   &\text{ if }$\tau = \alpha\to \beta$\\
\tau   &\text{ otherwise.}
 \end{cases}$ 
 \STATE\hspace{0cm} argTypes$(\tau) = \begin{cases}
[\alpha] + \text{argTypes}(\beta)   &\text{ if }$\tau = \alpha\to \beta$\\
[]   &\text{ otherwise.}
 \end{cases}$
\end{algorithmic}
\end{algorithm}

@@

\begin{figure*}
\centering  \begin{tabular}{lll}
    \toprule
    Domain&Example programs in frontiers&Example proposed subexpression\\\midrule
    Circuits&
    \begin{tabular}{l}
      \texttt{(lambda x (nand x x))}\\
      \texttt{(lambda x (lambda y (nand x (nand y y))}
    \end{tabular}&
    \texttt{(nand z z)}\\\midrule
    Strings&
    \begin{tabular}{l}
      \texttt{(lambda s (+ ',' (index 0 (split ',' s))))}\\
      \texttt{(lambda s (index 0 (split ' ' s)))}
    \end{tabular}&
    \texttt{(index 0 (split z s))}\\\bottomrule
  \end{tabular}
  \caption{The DSL induction algorithm works by proposing subexpressions of programs to add to the DSL.
    These subexpressions are taken from programs in the frontiers (middle column), and can introduce new variables (\texttt{z} and \texttt{s} in the right column).} 
  \label{fragmentExample}
\end{figure*}

@@

What DSL does \system learn?
The learned DSL contains templates for polynomials of different orders,
which lets the algorithm quickly hone in on the kinds of functions that are most appropriate to this domain (Table~\ref{regressionDSL}).
Examining the programs themselves,
one finds that the algorithm discovers representations for each of the polynomials that minimizes the number of continuous degrees of freedom:
for example, it represents the linear function $-x+2$ with the program 
\texttt{(* (* (+ x }$\mathcal{R}$\texttt{) }$\mathcal{R}$\texttt{) x)}, which has two continuous degrees of freedom -- while representing quartic functions using the invented DSL primitive $f_6$ in Table~\ref{regressionDSL}.
This falls out of our Bayesian treatment of the program learning problem -- both the bias towards shorter programs (more likely under $\mathcal{D},\theta$) and the likelihood model's BIC penalty.
\begin{table}\centering
\begin{tabular}{l}
  \toprule
  %% Primitives& $+,\times:\mathbb{R}\to \mathbb{R}\to \mathbb{R}$\\
  %% & $\mathcal{R}:\mathbb{R}$ (real valued parameter)\\\midrule
  %% Observation $x$& $N$ input/output examples: $\left\{(i_n,o_n) \right\}_{n\leq N}$\\
  %% \midrule
  %% Likelihood $\probability[x|p]$& $\propto\exp\left( - D\log N \right)\prod_{n\leq N}\mathcal{N}(p(i_n)|o_n)$
  %% \\\midrule
    $f_0 = $\texttt{(+ }$\mathcal{R}$\texttt{)}\\
    $f_1($\texttt{x}$) =$ \texttt{(* (* (+ x }$\mathcal{R}$\texttt{) }$\mathcal{R}$\texttt{) x)} \\
      $f_2($\texttt{x}$) =$ \texttt{(+ }$\mathcal{R}$\texttt{ ($f_1$ x))} \\
      $f_3($\texttt{x}$) = $\texttt{(* x (}$f_2$\texttt{ x))}\\
  $f_4($\texttt{x}$) = $\texttt{(}$f_0$\texttt{ (}$f_3$\texttt{ x)))}\\
  $f_5($\texttt{x}$) =$ \texttt{(* x (}$f_4$\texttt{ x))}\\
  $f_6($\texttt{x}$) = $\texttt{(}$f_0$\texttt{ (}$f_4$\texttt{ x))} (\emph{4th order polynomial})
  \\\bottomrule
\end{tabular}
\caption{Some induced DSL primitives for symbolic regression. System starts out with addition, multiplication, and real numbers, and learns functions for building 4th order polynomials.}\label{regressionDSL}
\end{table}

@@

\subsection{String editing}\label{textSection}
Synthesizing programs that manipulate strings is a classic problem in the
programming languages and AI literatures~\cite{menon2013machine,lau2001programming},
and algorithms that learn string editing programs ship in Microsoft Excel~\cite{gulwani2011automating}.
However, this prior work presumes a ready-made DSL,
expertly crafted to suit string editing.
We show in this experiment that we can instead start out with a
less specialized set of primitives -- close to the string routines provided by Python --
and automatically recover many of the higher-level building blocks that have made these
other system successful.
Table~\ref{textPrimitives} lists the primitive functions provided to \system.
Absent are any routines for searching for characters, extracting substrings,
or replacing one character with another.
Instead, we provide array/string indexing primitives (\texttt{slice}, \texttt{nth}),
basic arithmetic, and functions for splitting a string into a list (\texttt{split})
or turning a list into a string (\texttt{join}).

@@

\begin{table}[t]
  \centering
  \begin{tabular}{c}
    String editing primitives\\
    \begin{tabular}{l}
    \toprule
    \texttt{0}, \texttt{+1},\texttt{-1}\\
    \texttt{split}, \texttt{join}, \texttt{map}, \texttt{length}, \texttt{slice}\\
    \texttt{append}, \texttt{chr->str}, \texttt{trim}, \texttt{nth}\\
    \texttt{lowercase}, \texttt{uppercase}, \texttt{capitalize}\\
    \texttt{""},    \texttt{","},     \texttt{" "},    \texttt{"<"}, \texttt{">"}  
  \\\bottomrule
    \end{tabular}\\
\\    Subset of learned DSL\\
    \begin{tabular}{l}
      \toprule
      $f_0($\texttt{c}$) = $\texttt{(+ (chr->str c))}\\
      $f_1($\texttt{i,c,s}$) = $      \texttt{(+ (nth i (split c s))}\\\hspace{2.55cm}\texttt{ (chr->str c))}\\
      $f_2($\texttt{s,c}$) = $\texttt{((}$f_0$\texttt{ c) (}$f_1$\texttt{ (1+ 0) c s))} \\
      \hspace{0.5cm}($f_2$: \emph{Extract substring delimited by character})\\
      $f_3($\texttt{c,s}$) = $\texttt{(nth (1+ 0) (split c (nth 0 s)))}\\
      \hspace{0.5cm}($f_3$: \emph{2nd substring delimited by character})
    %% $f_0($\texttt{i,c,s}$) = $    \texttt{(+ (index i (split c s)) (chr->str c))}\\
    %% $f_1 = $\texttt{(join "")}
  \\\bottomrule
    \end{tabular}
  \end{tabular}
  \caption{Top: provided string editing primitives. Bottom: some induced DSL primitives.}\label{textPrimitives}
  \end{table}
%% \begin{figure}

@@

  \begin{minipage}{7.5cm}
 $f($\texttt{s}$) = $   \texttt{(}$f_3$\texttt{ (split '>' (upper s)) '<')}
    \end{minipage}
  \caption{Example string edit task (top) and program  \system writes (bottom). $f_3$ is a new DSL subroutine discovered by \system, defined in Table~\ref{textPrimitives}}\label{exampleTextProblem}
  \end{figure}

@@

\begin{table}
\centering
\begin{tabular}{| l | l | l |}
  \hline
  \emph{name} & \emph{input} & \emph{output} \\
  \hline
  add-3 & [1\, 2\, 3\, 4] & [4\, 5\, 6\, 7] \\
  append-4 & [7\, 0\, 2] & [7\, 0\, 2\, 4] \\
  len & [3\, 5\, 12\, 1] & 4 \\
  range & 3 & [1\, 2\, 3] \\
  has-2 & [4\, 5\, 7\, 4] & \texttt{false} \\
  has-head-in-tail & [4\, 5\, 7\, 4] & \texttt{true} \\
  repeat-2 & [7\, 0] & [7\, 0\, 7\, 0] \\
  drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
  \hline
\end{tabular}
\caption{Examples of tasks from our list function domain. These are simpler
  than most other tasks in our dataset which are too complex to warrant
  description here.}
\label{listexamples}
\end{table}

We take approaches from two different initial DSLs, \emph{base} and
\emph{rich}, and assess their performance. The base DSL starts with only
low-level primitives such as \texttt{mapi}, \texttt{reducei}, and
\texttt{if}, whereas the rich DSL includes some common structure that can be
built from these such as \texttt{all}, \texttt{filter}, and \texttt{slice}.
Both are theoretically capable of solving all tasks supplied by our dataset.
See Supplement for more detail on these DSLs.

>>>>

\newcommand{\system}{\textsc{DreamCoder}~}
\newcommand{\lowerBound}{\mathscr{L}}
\newcommand{\code}[1]{{\footnotesize\texttt{#1}}}
% Use the following line for the initial blind version submitted for review:
\usepackage[nohyperref]{icml2018}

@@

We apply \system to four domains:
 building Boolean circuits; symbolic regression; FlashFill-style~\cite{gulwani2011automating} string processing problems; and Lisp-style functions on lists.
 For each of these we deliberately provide an impoverished
 set of programming primitives,
 and show that our algorithm discovers
 its own domain-specific vocabulary for expressing solutions the domain (Table~\ref{initialExampleDSL}).
 \begin{table}
    \begin{tabular}{ll}
   \toprule
   Domain&Part of the learned DSL\\\midrule
   Circuits&\code{(nand (nand a a) (nand b b))}\\
   &\hspace{0.5cm} \emph{(an OR gate)}\\
   Regression& \code{(+ (* real x) real)} \\
      &\hspace{0.5cm} \emph{(a linear function of x)}\\
   Strings& \code{(join '' (split c s))}\\
   &\hspace{0.5cm} \emph{(delete occurrences of c in string s)}\\
   Lists& \code{(map (lambda (x) (+ x a)) l)}\\
   &\hspace{0.5cm} \emph{(add a to every element of list l)}
   \\\bottomrule
    \end{tabular}
    \caption{ Examples of structure found in DSLs  learned by our algorithm. \system builds a new DSL by discovering and reusing useful subroutines.}\label{initialExampleDSL}
 \end{table}

@@

To make this concrete we need to define what programs actually are and
what form $\probability[p |\mathcal{D},\theta]$ takes.
We represent programs as $\lambda$-calculus expressions.
$\lambda$-calculus is a formalism for expressing functional programs
that closely resembles the Lisp programming language.
$\lambda$-calculus includes variables, function application, and the ability to create new functions.
Throughout this paper we will write $\lambda$-calculus expressions in Lisp syntax.
Our programs are all strongly typed.
We use the Hindley-Milner polymorphic typing system~\cite{pierce} which is used in functional languages like OCaml.
Variables that range over types are always written using lowercase Greek letters
and we write $\alpha\to \beta$ to mean a function that takes as input type $\alpha$
and returns something of type $\beta$.
We use the notation $p:\tau$ to mean that the $\lambda$-calculus expression $p$
has the type $\tau$.
For example, to describe the type of the identity function
we would say \code{(lambda (x) x)}$:\alpha\to \alpha$.
We say a type $\alpha$ \emph{unifies} with $\tau$ if every expression
$p:\alpha$ also satisfies $p:\tau$. Furthermore, the act of \emph{unifying}
a type $\alpha$ with $\tau$ is to introduce constraints on the type
variables of $\alpha$ to ensure that $\alpha$ unifies with $\tau$.
See Supplement for more detail on program representation.
With this notation in hand we can now define DSLs:

@@

\begin{algorithm}[tb]
   \caption{Generative model over programs}
   \label{programGenerativeModel}
   \begin{algorithmic}
     \STATE \textbf{function} sampleProgramFromDSL$(\mathcal{D}, \theta, \tau)$:
  \STATE {\bfseries Input:} DSL $\mathcal{D}$, weight vector $\theta$, type $\tau$
  \STATE \textbf{Output:} a program whose type unifies with $\tau$
  \STATE \textbf{return} sample$(\mathcal{D}, \theta, \varnothing, \tau)$
\STATE
     \STATE \textbf{function} sample$(\mathcal{D}, \theta, \mathcal{E}, \tau)$:
  \STATE {\bfseries Input:} DSL $\mathcal{D}$, weight vector $\theta$, environment $\mathcal{E}$, type $\tau$
  \STATE \textbf{Output:} a program whose type unifies with $\tau$
  \IF{$\tau = \alpha\to\beta$}
  \STATE var $\gets$ an unused variable name
  \STATE body $\sim$ sample$(\mathcal{D},\theta,\{\text{var}:\alpha\}\cup\mathcal{E},\beta)$
   \STATE \textbf{return} \code{(lambda (}var\code{) }body\code{)}
   \ENDIF
   %   \ELSE
   \STATE $\text{primitives} \gets\{p | &p: \tau' \in \mathcal{D}\cup\mathcal{E}$
   \STATE \hspace{2.5cm}$\text{if }\tau\text{ can unify with yield}(\tau') \} $

   \STATE Draw $e\sim \text{primitives}$, w.p. $\propto\theta_e$ if $e\in \mathcal{D}$
   \STATE \hspace{3.1cm}w.p. $\propto\frac{\theta_{var}}{|\text{variables}|}$ if $e\in \mathcal{E}$
   \STATE Unify $\tau$ with yield$(\tau')$.
   \STATE $\left\{\alpha_k \right\}_{k\leq K}\gets\text{argTypes}(\tau')$
%   \STATE unify$(\tau,\beta)$
   \FOR{$k=1$ {\bfseries to} $K$}
 \STATE $a_k\sim\text{sample}(\mathcal{D},\theta,\mathcal{E},\alpha_k)$
 \ENDFOR
 \STATE \textbf{return} \code{(}$e\;a_1\; a_2\; \cdots\; a_K$\code{)}
 \STATE\textbf{where:}
 \STATE  yield$(\tau) = \begin{cases}
\text{yield}(\beta)   &\text{ if }$\tau = \alpha\to \beta$\\
\tau   &\text{ otherwise.}
 \end{cases}$ 
 \STATE\hspace{0cm} argTypes$(\tau) = \begin{cases}
[\alpha] + \text{argTypes}(\beta)   &\text{ if }$\tau = \alpha\to \beta$\\
[]   &\text{ otherwise.}
 \end{cases}$
\end{algorithmic}
\end{algorithm}

@@

\begin{figure*}
\centering  \begin{tabular}{lll}
    \toprule
    Domain&Example programs in frontiers&Example proposed subexpression\\\midrule
    Circuits&
    \begin{tabular}{l}
      \code{(lambda x (nand x x))}\\
      \code{(lambda x (lambda y (nand x (nand y y))}
    \end{tabular}&
    \code{(nand z z)}\\\midrule
    Strings&
    \begin{tabular}{l}
      \code{(lambda s (+ ',' (index 0 (split ',' s))))}\\
      \code{(lambda s (index 0 (split ' ' s)))}
    \end{tabular}&
    \code{(index 0 (split z s))}\\\bottomrule
  \end{tabular}
  \caption{The DSL induction algorithm works by proposing subexpressions of programs to add to the DSL.
    These subexpressions are taken from programs in the frontiers (middle column), and can introduce new variables (\code{z} and \code{s} in the right column).} 
  \label{fragmentExample}
\end{figure*}

@@

What DSL does \system learn?
The learned DSL contains templates for polynomials of different orders,
which lets the algorithm quickly hone in on the kinds of functions that are most appropriate to this domain (Table~\ref{regressionDSL}).
Examining the programs themselves,
one finds that the algorithm discovers representations for each of the polynomials that minimizes the number of continuous degrees of freedom:
for example, it represents the linear function $-x+2$ with the program 
\code{(* (* (+ x }$\mathcal{R}$\code{) }$\mathcal{R}$\code{) x)}, which has two continuous degrees of freedom -- while representing quartic functions using the invented DSL primitive $f_6$ in Table~\ref{regressionDSL}.
This falls out of our Bayesian treatment of the program learning problem -- both the bias towards shorter programs (more likely under $\mathcal{D},\theta$) and the likelihood model's BIC penalty.
\begin{table}\centering
\begin{tabular}{l}
  \toprule
  %% Primitives& $+,\times:\mathbb{R}\to \mathbb{R}\to \mathbb{R}$\\
  %% & $\mathcal{R}:\mathbb{R}$ (real valued parameter)\\\midrule
  %% Observation $x$& $N$ input/output examples: $\left\{(i_n,o_n) \right\}_{n\leq N}$\\
  %% \midrule
  %% Likelihood $\probability[x|p]$& $\propto\exp\left( - D\log N \right)\prod_{n\leq N}\mathcal{N}(p(i_n)|o_n)$
  %% \\\midrule
    $f_0 = $\code{(+ }$\mathcal{R}$\code{)}\\
    $f_1($\code{x}$) =$ \code{(* (* (+ x }$\mathcal{R}$\code{) }$\mathcal{R}$\code{) x)} \\
      $f_2($\code{x}$) =$ \code{(+ }$\mathcal{R}$\code{ ($f_1$ x))} \\
      $f_3($\code{x}$) = $\code{(* x (}$f_2$\code{ x))}\\
  $f_4($\code{x}$) = $\code{(}$f_0$\code{ (}$f_3$\code{ x)))}\\
  $f_5($\code{x}$) =$ \code{(* x (}$f_4$\code{ x))}\\
  $f_6($\code{x}$) = $\code{(}$f_0$\code{ (}$f_4$\code{ x))} (\emph{4th order polynomial})
  \\\bottomrule
\end{tabular}
\caption{Some induced DSL primitives for symbolic regression. System starts out with addition, multiplication, and real numbers, and learns functions for building 4th order polynomials.}\label{regressionDSL}
\end{table}

@@

\subsection{String editing}\label{textSection}
Synthesizing programs that manipulate strings is a classic problem in the
programming languages and AI literatures~\cite{menon2013machine,lau2001programming},
and algorithms that learn string editing programs ship in Microsoft Excel~\cite{gulwani2011automating}.
However, this prior work presumes a ready-made DSL,
expertly crafted to suit string editing.
We show in this experiment that we can instead start out with a
less specialized set of primitives -- close to the string routines provided by Python --
and automatically recover many of the higher-level building blocks that have made these
other system successful.
Table~\ref{textPrimitives} lists the primitive functions provided to \system.
Absent are any routines for searching for characters, extracting substrings,
or replacing one character with another.
Instead, we provide array/string indexing primitives (\code{slice}, \code{nth}),
basic arithmetic, and functions for splitting a string into a list (\code{split})
or turning a list into a string (\code{join}).

@@

\begin{table}[t]
  \centering
  \begin{tabular}{c}
    String editing primitives\\
    \begin{tabular}{l}
    \toprule
    \code{0}, \code{+1},\code{-1}\\
    \code{split}, \code{join}, \code{map}, \code{length}, \code{slice}\\
    \code{append}, \code{chr->str}, \code{trim}, \code{nth}\\
    \code{lowercase}, \code{uppercase}, \code{capitalize}\\
    \code{""},    \code{","},     \code{" "},    \code{"<"}, \code{">"}  
  \\\bottomrule
    \end{tabular}\\
\\    Subset of learned DSL\\
    \begin{tabular}{l}
      \toprule
      $f_0($\code{c}$) = $\code{(+ (chr->str c))}\\
      $f_1($\code{i,c,s}$) = $      \code{(+ (nth i (split c s))}\\\hspace{2.55cm}\code{ (chr->str c))}\\
      $f_2($\code{s,c}$) = $\code{((}$f_0$\code{ c) (}$f_1$\code{ (1+ 0) c s))} \\
      \hspace{0.5cm}($f_2$: \emph{Extract substring delimited by character})\\
      $f_3($\code{c,s}$) = $\code{(nth (1+ 0) (split c (nth 0 s)))}\\
      \hspace{0.5cm}($f_3$: \emph{2nd substring delimited by character})
    %% $f_0($\code{i,c,s}$) = $    \code{(+ (index i (split c s)) (chr->str c))}\\
    %% $f_1 = $\code{(join "")}
  \\\bottomrule
    \end{tabular}
  \end{tabular}
  \caption{Top: provided string editing primitives. Bottom: some induced DSL primitives.}\label{textPrimitives}
  \end{table}
%% \begin{figure}

@@

  \begin{minipage}{7.5cm}
 $f($\code{s}$) = $   \code{(}$f_3$\code{ (split '>' (upper s)) '<')}
    \end{minipage}
  \caption{Example string edit task (top) and program  \system writes (bottom). $f_3$ is a new DSL subroutine discovered by \system, defined in Table~\ref{textPrimitives}}\label{exampleTextProblem}
  \end{figure}

@@

\begin{table}
\centering
\begin{tabular}{| l | l | l |}
  \hline
  \emph{name} & \emph{input} & \emph{output} \\
  \hline
  add-3 & [1\, 2\, 3\, 4] & [4\, 5\, 6\, 7] \\
  append-4 & [7\, 0\, 2] & [7\, 0\, 2\, 4] \\
  len & [3\, 5\, 12\, 1] & 4 \\
  range & 3 & [1\, 2\, 3] \\
  has-2 & [4\, 5\, 7\, 4] & \code{false} \\
  has-head-in-tail & [4\, 5\, 7\, 4] & \code{true} \\
  repeat-2 & [7\, 0] & [7\, 0\, 7\, 0] \\
  drop-3 & [0\, 3\, 8\, 6\, 4] & [6\, 4] \\
  \hline
\end{tabular}
\caption{Examples of tasks from our list function domain. These are simpler
  than most other tasks in our dataset which are too complex to warrant
  description here.}
\label{listexamples}
\end{table}

We take approaches from two different initial DSLs, \emph{base} and
\emph{rich}, and assess their performance. The base DSL starts with only
low-level primitives such as \code{mapi}, \code{reducei}, and
\code{if}, whereas the rich DSL includes some common structure that can be
built from these such as \code{all}, \code{filter}, and \code{slice}.
Both are theoretically capable of solving all tasks supplied by our dataset.
See Supplement for more detail on these DSLs.
